---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

The code chunk appears:
```{r}
animals<-c("Snake","Bird","Dog","Spider")
animals
```

Type any R code in the chunk, for example:
```{r}
num_legs<-c(0,2,4,8)
num_legs
```
```{r}
animals_df<-data.frame(animals,num_legs)
animals_df
```
```{r}
ls()
```

## Matrix operations

**Use the seq() function to generate a sequence of numbers starting at 12 and decreasing to 2 in
steps of -2. Call this vector x vect. You may want to run ?seq or help(seq) to help you do
this**
```{r}
x_vect<-seq(12,2,-2)
x_vect
```

**Now convert the vector x vect into a matrix (with 2 rows and 3 coloumns) called X, using
the matrix() function. The result should look like this:**
```{r}
X<-matrix(x_vect,2,3)
X
```

**Next create a 2 by 2 matrix called Y consisting of a sequence of four numbers from 1-4. The
matrix Y should look like this:**
```{r}
Y<-matrix(seq(4),2,2)
Y
```

**In addition, create another 2 by 2 matrix called Z which looks as follows:**
```{r}
Z<-matrix(seq(4,10,2),2,2)
Z
```

**1. Use the t() function to compute Y^^T and Z^^T**
**2. Matrix multiplication %*%**

```{r}
result1<-Y%*%Z
result2<-Z%*%Y
result1
result2
```


```{r}
result3<-Y%*%X
result3
```

**3. Matrix element-wise multiplication**
```{r}
result4<-Y*Z
result4
```

**4. Matrix inverse**
get the result of Y^^-1
```{r}
M_inverse<-solve(Y)
M_inverse
```
```{r}
M_YY<-M_inverse%*%Y
M_YY
```
```{r}
M_YX<-M_inverse%*%X
M_YX
```

### about solve()
use ?solve() to check the usage of solve()

solve(Y,X) means $Y^{-1}\cdot X$
```{r}
result<-solve(Y,X)
result
```

